Authors Response to Reviewer's Comments

### 2023-10-03 manuscript/review/authorsResponse.qmd
changes following the review

date: last-modified
link-citations: true
linkcolor: blue
number-sections: true
format: docx
---

## Internal instructions

@@

## Journal formatting (ToDo list)

@@

## Response to the Editor

### General Comments

Dear Dr. Díaz-Yáñez:

Thank you very much for submitting your manuscript "Tree regeneration in models of forest dynamics: a key priority for further research" for review by Ecosphere. The reviewers and I appreciate the work you have accomplished. I am willing to consider a revised version for publication in the journal, assuming that you are able to modify the manuscript according to the reviewers' extensive recommendations. Your revisions should address the specific points made in the review comments.  Please note that Reviewer #2 made extensive comments directly on the manuscript.\
I should note that personally I can hardly disagree with your basic conclusion that models of forest dynamics need to do a good job of capturing regeneration and recruitment processes.  I have devoted most of my research over the past 40 years to developing the field methods to allow incorporation of accurate seedling, sapling and canopy recruitment in the models SORTIE and SORTIE-ND, in systems ranging from the tropics to boreal forests. The specific ecological processes incorporated in those models depended on the goals of the modeling and the ecology of the system, and included processes such as masting cycles, effects of disturbance on seedbed substrate dynamics, seed and seedling predation by small mammals, browsing by ungulates, Janzen-Connell effects, and, of course, detailed analyses of shading and neighborhood competition. More recently, this includes developing statistical models of seedling recruitment from the US forest inventory network that have been directly incorporated in the model for forests of the entire eastern US.  To my mind, the difficulties the models you reviewed encountered in capturing canopy recruitment speak more to the limitations of the approaches used in model development than in any inherent difficulties in capturing regeneration and recruitment in models of forest dynamics.

Sincerely,

Dr. Charles Canham Subject-matter Editor, Ecosphere

### Response to editors comments

Dear Dr. Charles Canham,

I want to express my gratitude for taking the time to review our manuscript. Your insightful thoughts on our findings were very helpful.

Our main takeaway from this manuscript is that while it may be possible to calibrate all 15 models against the EuFoRIa data, it is important to ensure that this aligns with the study's goals. If the goal is to provide extrapolation into a future climate, calibrating against data under the current climate may not be reliable. This is why we believe it is valuable to test model performance without any changes or calibration as we have done in this study.

We have carefully reviewed and addressed each of the specific points made in the review comments in this revised version. We believe that the changes we made have significantly improved the manuscript. Once again, I want to thank you for the time and effort you dedicated to reviewing our work. We look forward to hearing from you again.

Best regards,

Olalla Díaz-Yáñez on behalf of all the co-authors

## Response to reviewer 1

### General Comments

@@

##### Our general comments to reviewer 1:

Dear Reviewer 1,

Thank you for taking the time to review our manuscript. We appreciate your thoughtful comments and have carefully considered each of them. We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of the work.

We understand your concern about the length of the manuscript and the number of figures included. We have made an effort to reduce the text where possible and have verified that we are following the journal guidelines. Despite our best efforts, however, it has been difficult to significantly reduce the length of the manuscript.

Please find below our responses to each of your comments and the concrete actions we have taken in response.

Thank you again for your valuable feedback.

### Major comments

#### [COMMENT 1]{.underline}

'regeneration' versus 'recruitment'. I personally have had this conversation multiple times over the years, so I understand that there is variation in the definitions and uses. However, the authors do define these terms in the methods (L167-171) and I completely agree with them. Regeneration is for flowers, seeds, and seedlings while recruitment is for reaching a certain size threshold. However, the entire manuscript refers to regeneration as "passing of trees across a specific diameter threshold ("ingrowth") (L142), either reaching a size of 7 or 10 cm DBH. This is recruitment. Since this is the response variable for the manuscript, I would suggest replacing 'regeneration' with 'recruitment' throughout the manuscript.\

##### Response to comment 1:

[We should discuss this point again... I do not agree to simple change regeneration with recruitment. Specially without a detailed definition as we had in the previous version.]{.underline}

[SUGGESTION FROM YANNEK:]{.underline}

[Thank you for pointing this out. We also discussed terminology several times and we agree. We followed your suggestion replaced 'regeneration' with 'recruitment' where the difference is not clear. However, we believe that in some places regeneration is more accurate. Especially in many parts of the introduction and discussion. We argue that regeneration is often only modeled via "recruitment", but with the aim to simulate the whole process of regeneration. In this context it makes more sense to use the general term (e.g., for discussion consequences.)]{.underline}

[COMMENT from YK: I would argue that recruitment is evidence for successful regeneration. This allows us to use recruitment for discussions observation vs. simulation in both aspects: technical (recruitment) and general (regeneration).]{.underline}

#### COMMENT 2

@@

##### Response to comment 2:

Thank you for bringing to our attention the issue with the way we communicated our diameter thresholds approach in the manuscript. The observed data had varying diameter thresholds for reporting tree recruitment, and, for the purposes of this study, we merged the recruited trees values into two thresholds of 7 and 10cm.

In total, we had 165 sites where the records showed the number of recruited trees at a 7cm threshold. For these sites, any recruited tree above 7cm was counted. We also had 35 sites where the minimum diameter threshold was 10cm, and for these sites, we could only account for trees recruited from 10cm diameter. We used the other 165 sites with 7cm diameter threshold in the 10cm database, but only accounting for newly established trees above 10cm. There was no like-to-like comparison for any other diameter threshold, only for 7 and 10cm.

The reviewer correctly pointed out that in Figure 2 for the 7cm threshold, there were 165 "dots" in the observed data, while for the 10cm diameter threshold, there were 200 "dots." However, simulated values were obtained for the 200 sites for both thresholds, as explained in the Simulation protocol section. Therefore, the simulated values in the figures are from the 200 sites.

When examining the change in mortality between 7 to 10cm or species diversity comparison between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10cm thresholds in the observed data. This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript.

To better clarify this aspect, we have made the following changes:

1\) We have better explained the difference on the observed data in the section 3.2 Observed data with the following text:

@@

2\) We have added the following text in the caption of figure 4

@@

#### [COMMENT 3]{.underline}

@@

##### Response to comment 3:

[I am going to do this.]{.underline}

### Minor comments:

#### COMMENT 4

@@

##### Response to comment 4:

We have made some changes to the abstract based on feedback from both reviewers, including clarifying certain parts for better understanding. The models were able to capture some aspects of regeneration more effectively than others. One example, is the sentence highlighted by the reviewer that was refereing to the diversity of regeneration.

#### COMMENT 5

@@

##### Response to comment 5:

Following the reviewers comments we have made several changes in this paragraph to improve its readability and flow if ideas with the main purpose of the study text.

#### COMMENT 6

Table 1. Please clarify what the letters are for (a, b, c and d in the table headers). For models that start from saplings or ingrowth, please add in the size as well (i.e., start from saplings with DBH = 1.0). Could also add in the classification scheme mentioned in the methods (L168, regeneration or recruitment) to Table 1. \

##### Response to comment 6:

Following the reviewers comment we have added a new column classifying the models in recruit or regeneration models; and the letters a,b,c, d,e and f table are now better described. Regarding the suggestion on adding into this table the size of the saplings or ingrowth: there are models that start from saplings or ingrowth what without a constant diameter threshold as they are biomass based (e.g. LandClim), we find it complicated to add all this information in this table without making it too large and therefore we have decided not to included. Furthermore the models reports in the supplementary materials provide more details on how the models simulate regeneration.

ToDo All models check that the new column in the table is correct for your model.

#### COMMENT 7

L218 -- do I understand this correctly, that regeneration rates are a mean of 56 trees per ha per decade? This seems really low (assuming that this is for \>10 cm DBH).  \

##### Response to comment 7:

Yes, this is the mean value across all observations and sites, there are two main reasons why this number might seem low, first due to the zero inflated nature of the data, and second because these are forest reserves, corresponding with denser forests in which regeneration would be expected to be lower. This value is in agreement with a further observed data from reserved (Käber, Y., et al (2023). Sheltered or suppressed? Tree regeneration in unmanaged European forests. Journal of Ecology, 00, 1--15. https://doi.org/10.1111/1365-2745.14181)

#### COMMENT 8

@@

##### Response to comment 8:

For those models that simulated more than the eleven species, we created a different category "others" were we lumped those extra recruited species. This were counted when we calculated the diversity index and for total recruitment values. Following the reviewer comments we have added this information in the simulation protocol section with the following text:

"Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category "others""

#### COMMENT 9

@@

##### Response to comment 9:

Following the reviewers comment we have modified this text to improve it readability with:

@@

#### COMMENT 10

L364: so please remove the 4C model from the figure as it is not comparable. Do you need to also remove other models that don't simulate the same number of species too?\

##### Response to comment 10:

Even thought it is not directly comparable, we still believe is interesting for the reader that we share the findings related to the 4C model and those that simulate more species that were lumped in the category others.

#### [COMMENT 11]{.underline}

@@

##### Response to comment 11:

@@

#### COMMENT 12

@@

##### Response to comment 12:

Following the reviewers comment we have increased the legend size, and now the difference between observed and simulated is visible. We have also explained in the caption the meaning of the black line:

@@

#### COMMENT 13

@@

##### Response to comment 13:

We agree with the reviewer that it was interesting to find that the models with feedback did not captured better the species diversity of the regeneration compared with those without, with the exeption of ForCEEPS. The Figure 4 shows three representative patterns observed across models using 6 of them as an example, and the Figure S3 provides the details for each model. Models with feedback are defined in table 1 and are ForCEEPS(f) , PICUS, SIBYLA, xComp, iLand, LandClim, Landis-II, TreeMig, LPJ-GUESS, and aDGVM2. We have addressed this point in the discussion section with the following text:

@@

#### COMMENT 14

@@

##### Response to comment 14:

Following the reviewers recommendation we have changed the axis labels in the appendix figures.

#### COMMENT 15

@@

##### Response to comment 15:

This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands and assumes a fixed relationship between the number of stems and the quadratic mean diameter in fully stocked pure stands.

#### COMMENT 16

@@

##### Response to comment 16:

Thank you! We also think is a good summary figure.

#### [COMMENT 17]{.underline}

@@

##### Response to comment 17:

@@

#### COMMENT 18

@@

##### Response to comment 18:

Following the reviewer comment we have removed figure S7.

#### COMMENT 19

@@

##### Response to comment 19:

Figure 4 shows three examples of the patterns observed across models (overestimation intermediate and underestimation) using one exemplary model of each category , the actual patterns per model can be found in Figure S3. In the results we showed that most models captured reasonably well or overpredicted species diversity at the stand level, as most of the models are under the patterns presented in plot 1 and 2 from A on Figure 4. We have decided to keep the referenced statement link to this findings in the discussion as it was.

#### COMMENT 20

@@

##### Response to comment 20:

In this statement we were referring only to regeneration levels. Although we agree with the reviewer that we did state that most of the models overestimate regeneration, here our point is that even-thought they did overestimate regeneration they did not deviate exceedingly, specially under the consideration that these models have never been confronted with this type of data.

#### COMMENT 21

@@

##### Response to comment 21:

With this statement we were referring to the multiple potential combinations of approaches of modeling tree regeneration. Based on the reviewers comment we have modified this for clarity.

#### COMMENT 22

@@

##### Response to comment 22:

Thank you for noticing this. We have now corrected this mistake.

#### COMMENT 23

@@

##### Response to comment 23:

We agree with the reviewer that ideally models should be able to capture broader patterns. With this sentence we just wan to explain that the stochasticity of this process may have partly explained the results on this regard.

#### COMMENT 24

@@

##### Response to comment 24:

The reviewer is correct that landscape level models were not significantly performing worst than others. Here we just wanted to discuss that maybe another spatial consideration will have allowed landscape level models to better represent regeneration processes as they are designed for that.

#### COMMENT 25

@@

##### Response to comment 25:

We agree with the reviewer, a coarse representation of water balance would not make it more sensitive to water deficits. And this was exactly our point in this part of the discussion where we wrote: "This may at least partially explain the unsatisfactory performance of many models along the drought axis (i.e., climatic water balance)."

#### [COMMENT 26]{.underline}

@@

##### Response to comment 26:

Following the reviewers comment we have expanded this aspect in the following paragraph XXXX

[Both reviewers asked to be more specific on the data we are missing, what vairables would you say without being too detailed? also linked to comment 59 from reviewer 2]{.underline}

## Response to reviewer 2

### General Comments

The paper evaluates whether forest dynamics models are successful at capturing the essence of regeneration processes, by comparing outputs from 15 simulation models with tree recruitment data obtained from 300 mature forests spread across central/eastern Europe.  Specifically, the models predicted how many trees of different species recruited each year (i.e. grew larger than 7-cm or 10-cm stem diameter) and indices derived from these numbers were compared with analogous indices derived from repeat-measured forest inventory plots.  Most of the simulation models performed poorly for most of the indices examined.  In particular, they tended to overestimate recruitment rates.

@@

##### Our general comments to reviewer 2:

Dear Reviewer 2,

We would like to express our sincere gratitude for all the detailed comments they provided on our manuscript. We have carefully considered all of their feedback and made the necessary changes to improve the manuscript. Thanks to your comments, we have improved our manuscript.

Thank you again for your valuable feedback.

### Major comments

#### [COMMENT 1]{.underline}

@@

##### Response to comment 1:

[I think we have proposed some ideas of how to do this better in the discussion. What else do you think we should add here?]{.underline}

#### [COMMENT 2]{.underline} {#sec-R2C2}

@@

##### Response to comment 2:

Following the reviewer comment we have added a "road map" at the end of the introduction.

#### [COMMENT 3]{.underline}

@@

##### Response to comment 3:

[Decide final terminology]{.underline}

#### COMMENT 4

@@

##### Response to comment 4:

The reviewer is correct that most of the people involved in this manuscript have a central European focus in their current work, and so did the study set up. We have included some relevant global or with a focus outside Europe references when relevant. e would be happy to consider some other relevant suggestions.

### Minor comments:

@@

#### COMMENT 5

@@

##### Response to comment 5:

Based on the reviewers comments we have modified this sentence to: "However, there is a need to evaluate their capacity to accurately represent tree regeneration"

#### COMMENT 6

@@

##### Response to comment 6:

Following the reviewers comment we have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration.

#### COMMENT 7

@@

##### Response to comment 7:

Following the reviewers comments we have changed "feature" with "present".

#### COMMENT 8

@@

##### Response to comment 8:

Following the reviewers comment, we have changed the word "were" with "have been".

#### [COMMENT 9]{.underline}

@@

##### Response to comment 9:

Following the reviewers comment, we have corrected the reference style.

#### COMMENT 10

@@

##### Response to comment 10:

Following the reviewers comment, we have changed this sentence to: "This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience."

#### COMMENT 11

@@

##### Response to comment 11:

Following the reviewers comment, we have changed "such as" with "including"

#### COMMENT 12

##### Response to comment 12:

Following the reviewers comment, we have changed "such as" with "including"

#### COMMENT 13

@@

##### Response to comment 13:

Following the reviewer's comment we have changed "it" with "them"

#### [COMMENT 14]{.underline}

@@

##### Response to comment 14:

It is indeed interesting to account with the increasing data vailablity to better understand tree regeneration processes. However, we believe that the main take away it is not that it would be impossible to calibrate all 15 models against e.g. the EuFoRIa data to better capture regeneration, but that the focus should be in the study goals. And when the goal is to provide extrapolation into a future climate, then calibration against data under current climate may be unreliable. With the approach we have followed in this study where we have asked the models to simulate forest dynamics and capture regeneration without statistical predetermination we can better assess their behavior and understand how their current model structures and inclusion of exclusion of certain ecological processes impact the outputs we are seeing.

[Following this and other comments from the editor, we have tried to capture this better in the discussion/conclussions section. Should we add something on this?]{.underline}

#### COMMENT 15

@@

##### Response to comment 15:

@@

#### [COMMENT 16]{.underline}

@@

##### Response to comment 16:

[Following the reviewer's comment we have added a reference]{.underline}

@@

#### [COMMENT 17]{.underline}

@@

##### Response to comment 17:

[Following the reviewer's comment we have added a reference]{.underline}

@@

#### **COMMENT 18**

@@

##### Response to comment 18:

We currently use the exciting Hanbury-brown 2022 review in several places across our manuscripts, following the reviewer's comment here and the general comment 2 in @sec-R2C2 we have also get inspired by their structure by creating a road map to help the reader to navigate the manuscript content.

#### **COMMENT 19**

@@

##### Response to comment 19:

@@

####

#### **COMMENT 20**

L150 "regeneration niche of multiple species"

L154 "traits"

##### Response to comment 20:

We are unsure why the reviewer highlighted those words

#### **COMMENT 21**

@@

##### Response to comment 21:

This statement relates to Bugmann and Seidl (2022) where complexity values were provided. The exact details of how we calculated this mean regeneration formulation complexity value are available in the software suplementary materials to this paper: \`\`\`figure.R\`\`\` line 706-750. This value represents the mean complexity of regeneration formulations per model.

#### [**COMMENT 22**]{.underline}

@@

##### Response to comment 22:

[Is there a way to explain this shortly??]{.underline}

#### **COMMENT 23**

@@

##### Response to comment 22:

@@

#### **COMMENT 24**

@@

##### Response to comment 24:

@@

#### **COMMENT 26**

@@

##### Response to comment 26:

@@

#### **COMMENT 27**

@@

##### Response to comment 27:

Following the reviewer's comment we have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect."

#### **COMMENT 28**

@@

##### Response to comment 28:

@@

#### **COMMENT 29**

@@

##### Response to comment 29:

@@

**COMMENT 30**

@@

##### Response to comment 30:

@@

#### **COMMENT 31**

@@

##### Response to comment 31:

Please, see our reply on comment 25 in here @sec-R2C25r

#### **COMMENT 32**

@@

##### Response to comment 32:

@@

#### **COMMENT 33**

@@

##### Response to comment 33:

@@

#### [**COMMENT 34**]{.underline}

@@

#####

##### Response to comment 34:

@@

[We have re-assessed the terms used across this manuscript XXX]{.underline}

####

**COMMENT 35**

@@

##### Response to comment 35:

@@

#### **COMMENT 36**

@@

##### Response to comment 36:

@@

#### **COMMENT 37**

L230-231 "That is, the models were run in "blind flight" mode." Remove

##### Response to comment 37:

@@

#### **COMMENT 38**

@@

##### Response to comment 38:

@@

####

@@

##### Response to comment 39:

@@

#### **COMMENT 40**

@@

##### Response to comment 40:

@@

#### **COMMENT 41**

@@

##### Response to comment 41:

@@

#### **COMMENT 42**

@@

##### Response to comment 42:

@@

#### [**COMMENT 43**]{.underline}

@@

##### Response to comment 43:

[Change if we change regeneration with recruit]{.underline}

####

#### **COMMENT 44**

@@

##### Response to comment 44:

@@

#### **COMMENT 45**

@@

##### Response to comment 45:

@@

#### **COMMENT 46**

@@

##### Response to comment 46:

>>>>

date: last-modified
link-citations: true
linkcolor: blue
number-sections: true
format: pdf
---

# Internal instructions

@@

# Journal formatting (ToDo list)

@@

# Response to the Editor

## General Comments

Your revisions should address the specific points made in the review comments.  Please note that Reviewer #2 made extensive comments directly on the manuscript.\
I should note that personally I can hardly disagree with your basic conclusion that models of forest dynamics need to do a good job of capturing regeneration and recruitment processes.  I have devoted most of my research over the past 40 years to developing the field methods to allow incorporation of accurate seedling, sapling and canopy recruitment in the models SORTIE and SORTIE-ND, in systems ranging from the tropics to boreal forests. The specific ecological processes incorporated in those models depended on the goals of the modeling and the ecology of the system, and included processes such as masting cycles, effects of disturbance on seedbed substrate dynamics, seed and seedling predation by small mammals, browsing by ungulates, Janzen-Connell effects, and, of course, detailed analyses of shading and neighborhood competition. More recently, this includes developing statistical models of seedling recruitment from the US forest inventory network that have been directly incorporated in the model for forests of the entire eastern US.  To my mind, the difficulties the models you reviewed encountered in capturing canopy recruitment speak more to the limitations of the approaches used in model development than in any inherent difficulties in capturing regeneration and recruitment in models of forest dynamics.

**Response to editors comments**

Our main take-away from this manuscript is that while it would certainly be possible to calibrate all 15 models against the EuFoRIa data, it is important to ensure that this aligns with the study's goals. If the goal is to provide extrapolation into a future climate, calibrating against data under the current climate may not be reliable, as shown by the two models in our set that *were* calibrated against such data (i.e., SIBYLA and xComp). This is why we believe it is valuable to test the performance of the models "as they are", i.e. as we have done in this study.

We have carefully reviewed and addressed each of the specific points made in the review comments in this revised version, as explained below.

# Response to reviewer 1

## General Comments

@@

Our general comments to reviewer 1:

Thank you for taking the time to review our manuscript. We appreciate your thoughtful comments and have carefully considered each of them. We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of our manuscript.

We appreciate the concern about the length of the manuscript and the number of figures included. We have made an effort to reduce the text where possible and have verified that we are following the journal guidelines. Despite our best efforts, however, it has been difficult to significantly reduce the length of the manuscript.

Our responses to each of the comments and the concrete actions we have taken are detailed below.

## Major comments

### COMMENT 1

'regeneration' versus 'recruitment'. I personally have had this conversation multiple times over the years, so I understand that there is variation in the definitions and uses. However, the authors do define these terms in the methods (L167-171) and I completely agree with them. Regeneration is for flowers, seeds, and seedlings while recruitment is for reaching a certain size threshold. However, the entire manuscript refers to regeneration as "passing of trees across a specific diameter threshold ("ingrowth") (L142), either reaching a size of 7 or 10 cm DBH. This is recruitment. Since this is the response variable for the manuscript, I would suggest replacing 'regeneration' with 'recruitment' throughout the manuscript.

**Response to comment 1:**

Thank you for pointing this out. We discussed terminology several times and basically agree. We have followed your suggestion and have replaced the erroneous use of 'regeneration' by 'recruitment' where we refer to ingrowth into a certain DBH class. However, in some places the term 'regeneration' is more accurate, especially in many parts of the Introduction and Discussion sections. 'Regeneration' is a broader term that starts from flowering and ends with ingrowth into a given size class; that is, recruitment is the result of successful regeneration. Thus, where we refer to the entire process of regeneration, it would not be appropriate to use 'recruitment', and therefore we maintain the former term where it is fitting.

### COMMENT 2

@@

**Response to comment 2:**

We realized that there is an issue with the way we communicated our diameter thresholds approach in the manuscript. The observed data had varying diameter thresholds for reporting tree recruitment, and, for the purposes of this study, we merged the recruited trees values into two thresholds of 7 and 10 cm.  That is, in all plots (*n* = 165) that had a calipering limit \<7 cm, only the successful recruitment events at 7 cm were considered; and analogously for plots (*n* = 35) with a callipering limit between 7 and 10 cm, where only recruitment at 10 cm was considered. Thus, the 7 cm dataset (*n* = 165) actually is a subset of the 10 cm dataset (*n* = 200), as all data from the 7-cm threshold could be re-calculated for the 10-cm threshold. Thus, there was no 'like-to-like' comparison for any other diameter threshold, only for 7 and 10cm.

When examining the change in mortality between 7 and 10 cm, or the difference in species diversity between observed and simulated data (Figure 4), we could only use the 165 sites that contained both pieces of information at the 7 and 10 cm thresholds in the observed data. This is reported in the supplementary materials software, script \`\`\`Figures.R\`\`\` line 536 and L284-349 and now better explained in the manuscript:

1\) We have better explained the difference on the observed data in the section 3.2 Observed data L\[XX-XX\]

@@

2\) We have added a new text in the caption of figure 4 L\[XX-XX\]

@@

### [COMMENT 3]{.underline}

@@

**Response to comment 3:**

[PENDING: I am going to do this.]{.underline}

## Minor comments:

### COMMENT 4

@@

Response to comment 4:

We have changed the abstract based on feedback from both reviewers, including clarifying certain parts for better understanding. The models were able to capture some aspects of regeneration more effectively than others. One example, is the sentence highlighted by the reviewer that was referring to the diversity of regeneration.

### COMMENT 5

@@

**Response to comment 5:**

We have made several changes in this paragraph to improve its readability and flow of ideas towards explaining the main purpose of the study.

### COMMENT 6

Table 1. Please clarify what the letters are for (a, b, c and d in the table headers). For models that start from saplings or ingrowth, please add in the size as well (i.e., start from saplings with DBH = 1.0). Could also add in the classification scheme mentioned in the methods (L168, regeneration or recruitment) to Table 1.

**Response to comment 6:**

We have added a new column classifying the models into recruitment or regeneration models; and the letters a,b,c, d,e and f in the table are now better described. Regarding the suggestion of adding the size of the saplings or ingrowth: we are concerned that this would be more confusing than helpful. For example, there are models that start from saplings without a diameter threshold, as they are biomass-based (e.g. LandClim), and providing all information for each model would make this table rather overloaded, without a clear benefit, as we believe. Therefore we have decided not to include this information. We would also like to note that the 15 model reports in the Supplementary Materials provide more details how regeneration is simulated.

### COMMENT 7

L218 -- do I understand this correctly, that regeneration rates are a mean of 56 trees per ha per decade? This seems really low (assuming that this is for \>10 cm DBH).

**Response to comment 7:**

Yes, this is the mean value across all observations and sites. There are two major reasons why this number appears low. First due to the zero-inflated nature of the data, and second because these are forest reserves, typically being characterized by denser forests in which regeneration can be expected to be low. This value is in agreement with further observed data from reserves (Käber, Y., et al (2023). Sheltered or suppressed? Tree regeneration in unmanaged European forests. Journal of Ecology, https://doi.org/10.1111/1365-2745.14181).

### COMMENT 8

@@

**Response to comment 8:**

For those models that simulated more than the eleven species, we created a different category "others" were we lumped those extra recruited species. These were counted when we calculated the diversity index as well as for total recruitment values. We have added this information in the simulation protocol section L\[XX-XX\].

"Two models included their standard set of species for the simulations, which is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species simulated beyond the eleven species were lumped in a category"others""

### COMMENT 9

@@

**Response to comment 9:**

Following the reviewers comment we have modified the text L\[XX-XX\]

@@

### COMMENT 10

L364: so please remove the 4C model from the figure as it is not comparable. Do you need to also remove other models that don't simulate the same number of species too?

**Response to comment 10:**

We acknowledge the point that the reviewer is making, but even though it is not directly comparable, we still believe to be insightful for the reader that we share the findings related to the 4C model. Note that in those models that simulate more species, these additional species were lumped in the category 'others' (as explained further above), thus with little impact on the comparison.

### [COMMENT 11]{.underline}

@@

**Response to comment 11:**

@@

### COMMENT 12

@@

**Response to comment 12:**

We have increased the font size of the legend, and now the difference between observed and simulated is visible. We have also explained the meaning of the black line in the caption L\[XX-XX\].

@@

### COMMENT 13

@@

**Response to comment 13:**

We agree that it is interesting that the models with feedback did not capture better the species diversity of the regeneration compared with those without, with the exeption of ForCEEPS. Figure 4 shows three representative patterns observed across models using 6 of the models as an example, and Figure S3 provides the details for each model. We believe it is not very informative to color the points from the models with feedback in Figure 4 as this will only color all the points from some of the panels and it is possible to identify the models with feedback in table 1 We have addressed this point in the discussion section L\[XX-XX\].

@@

### COMMENT 14

@@

**Response to comment 14:**

Done.

### COMMENT 15

@@

**Response to comment 15:**

This paragraph is addressing the comparison with the theoretical Reineke value which is usually calculated for even-aged, single species stands. We have change this text to clarify this aspect.

### COMMENT 16

@@

**Response to comment 16:**

Thank you!

### [COMMENT 17]{.underline}

@@

**Response to comment 17:**

@@

### COMMENT 18

@@

**Response to comment 18:**

We have removed figure S7.

### COMMENT 19

@@

**Response to comment 19:**

Figure 4 shows three examples of the patterns observed across models (overestimation intermediate and underestimation) using one exemplary model of each category. The actual patterns per model can be found in Figure S3. In the results, we showed that most models capture reasonably well or overpredicted species diversity at the stand level, as most of the models are following the patterns presented in plot 1 and 2 (part A of Figure 4). We have decided to keep the referenced statement regarding this finding in the discussion.

### COMMENT 20

@@

Response to comment 20:

In this statement we were referring only to regeneration levels. Although we agree with the reviewer that we did state that most of the models overestimate regeneration, here our point is that even though they did overestimate regeneration, they did not deviate exceedingly, especially given that they have never been confronted with this type of data.

### COMMENT 21

@@

Response to comment 21:

With this statement, we were referring to the multiple potential combinations of approaches of modeling tree regeneration. We have modified this for clarity L\[XX-XX\]

### COMMENT 22

@@

**Response to comment 22:**

This was a glitch, which now is corrected L\[ XX-XX\]

### COMMENT 23

@@

**Response to comment 23:**

We have removed these statements.

### COMMENT 24

@@

Response to comment 24:

Yes, the landscape level models were not performing significantly worse than the others. However, they were not performing better either, potentially because they were used in "single site" rather than "landscape" mode. Thus, the explicit consideration of horizontal space would possibly allow the landscape level models to better represent regeneration processes, as they are designed for that. This, however, was completely out of the scope of the present analysis, as there are no landscape-level data of the surroundings of the 200 plots.

### COMMENT 25

@@

**Response to comment 25:**

Our point here is that the models, although many of them have detailed schemes for the water balance, may not have performed well because of the absence of site-specific (truly reliable) soil information.

### COMMENT 26

@@

**Response to comment 26:**

We have changed this text L\[XX-XX\]. Although there are multiple data sources (such as MASTIF, MASTREE, National Forest Inventories, Forest Reserve plots, etc.), usually these data do not refer to the same locations and thus cannot be used in a site-specific manner for calibrating or testing models that intend to follow the ‚regeneration' chain of processes.

"Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration."

# Response to reviewer 2

## General Comments

@@

**Our general comments to reviewer 2:**

We would like to thank Reviewer #2 for this positive overall assessment and the level of detail of the review. We have taken care to go through all suggestions made in the PDF, and have changed/improved the manuscript in most instances; however these suggestions are too numerous to be dealt with on an individual basis here.

## Major comments

### COMMENT 1

@@

Response to comment 1:

In this first assessment we were not aiming at identify what specific subcomponents may hold the key and therefore dased on the current approach we could not provide more concrete recommendation for model development which most likely be specific for each of the 15 models. Maybe we need more such studies, including calibration as a first step, and then application against a withheld part of the data; or application (and comparison) under climate change scenarios to evaluate the robustness of the signal under future conditions.

### COMMENT 2 {#sec-R2C2}

@@

**Response to comment 2:**

We have added a "road map" at the end of the introduction.

### [COMMENT 3]{.underline}

@@

**Response to comment 3:**

Thank you for pointing this out. We discussed terminology several times and basically agree. We have followed your suggestion and have now use recruitment to discuss the process and regeneration as the overarching term.

### COMMENT 4

@@

**Response to comment 4:**

Indeed, most of the people involved in this manuscript have a central European focus in their current work, and so did the study setup. We have included global references or references with a focus outside Europe when they were relevant. We have evaluated whether this could be further "internationalized", but felt our choices were appropriate.

## Minor comments:

@@

### COMMENT 5

@@

Response to comment 5:

We have modified this sentence L\[XX-XX\]

"However, there is a need to evaluate their capacity to accurately represent tree regeneration"

### COMMENT 6

@@

Response to comment 6:

We have changed the word "comprehensive" to "extensive" to express that we used a wide-range data that considers many important elements of regeneration.

### COMMENT 7

@@

**Response to comment 7:**

We have changed "feature" with "present".

### COMMENT 8

@@

**Response to comment 8:**

We have changed the word "were" with "have been".

### COMMENT 9

@@

**Response to comment 9:**

We have corrected the reference style.

### COMMENT 10

@@

**Response to comment 10:**

We have changed this sentence to: "This presents an important research gap, particularly in the context of climate-induced forest disturbances and forest resilience."

### COMMENT 11

@@

Response to comment 11:

We have changed "such as" with "including"

### COMMENT 12

@@

Response to comment 12:

We have changed "it" with "them"

### [COMMENT 13]{.underline}

@@

Response to comment 13:

It is indeed interesting to account with the increasing data availability to better understand tree regeneration processes. However, we believe that the main take away it is not that it would be impossible to calibrate all 15 models against e.g. the EuFoRIa data to better capture regeneration, but that the focus should be in the study goals. And when the goal is to provide extrapolation into a future climate, then calibration against data under current climate may be unreliable. With the approach we have followed in this study where we have asked the models to simulate forest dynamics and capture regeneration without statistical predetermination we can better assess their behavior and understand how their current model structures and inclusion of exclusion of certain ecological processes impact the outputs we are seeing.

### COMMENT 14

@@

Response to comment 14:

@@

### [COMMENT 15]{.underline}

@@

**Response to comment 15:**

@@

### [COMMENT 16]{.underline}

@@

**Response to comment 16:**

@@

### **COMMENT 17**

@@

**Response to comment 17:**

We currently use the exciting Hanbury-brown 2022 review in several places across our manuscripts, following the reviewer's comment here and the general comment 2 we have also get inspired by their structure by creating a road map to help the reader to navigate the manuscript content.

### **COMMENT 18**

@@

**Response to comment 18:**

@@

### **COMMENT 19**

@@

Response to comment 19:

This statement relates to Bugmann and Seidl (2022) where complexity values were provided. The exact details of how we calculated this mean regeneration formulation complexity value are available in the software supplementary materials to this paper: \`\`\`figure.R\`\`\` line 706-750. This value represents the mean complexity of regeneration formulations per model.

### [**COMMENT 20**]{.underline}

@@

Response to comment 22:

[Is there a way to explain this shortly \@ Yannek ??]{.underline}

### **COMMENT 23**

@@

Response to comment 22:

@@

### **COMMENT 24**

@@

Response to comment 24:

@@

### **COMMENT 26**

@@

Response to comment 26:

@@

### **COMMENT 27**

@@

Response to comment 27:

We have changed "and exposition (i.e.,slope and aspect)." with ", slope and aspect."

### **COMMENT 28**

@@

Response to comment 28:

@@

### **COMMENT 29**

@@

Response to comment 29:

@@

### **COMMENT 30**

@@

Response to comment 30:

@@

### **COMMENT 31**

@@

Response to comment 31:

Please, see our reply on comment 25

### **COMMENT 32**

@@

Response to comment 32:

@@

### **COMMENT 33**

@@

Response to comment 33:

@@

### [**COMMENT 34**]{.underline}

@@

Response to comment 34:

@@

We have re-assessed the terms used across this manuscript and change it across the text accordingly.

### **COMMENT 35**

@@

Response to comment 35:

@@

### **COMMENT 36**

@@

Response to comment 36:

@@

### **COMMENT 37**

L230-231 "That is, the models were run in"blind flight" mode." Remove

Response to comment 37:

@@

### **COMMENT 38**

@@

Response to comment 38:

@@

@@

Response to comment 39:

@@

### **COMMENT 40**

@@

Response to comment 40:

@@

### **COMMENT 41**

@@

Response to comment 41:

@@

### **COMMENT 42**

@@

Response to comment 42:

@@

### **COMMENT 43**

@@

Response to comment 43:

We have change the terminology.

### **COMMENT 44**

@@

Response to comment 44:

@@

### **COMMENT 45**

@@

Response to comment 45:

@@

### **COMMENT 46**

@@

Response to comment 46:
